---
title: "PS270 Final Project"
author: "Miles Vogel"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction:

I am interested in exploring data related to...

the immense changes we have witnessed occuring within the Republican party since the rise of Donald Trump. I would like to use various figures and statistical evidence to show numerous ways in which the Republican base has shifted since Donald Trump's rise to prominence during the 2016 election cycle. I want to analyze whether Trump's rise has fostered an increase in hateful rhetoric in the United States.

# Research Question: 

To what extent has Donald Trump and the Maga-Era of Republican politics changed the compositions and backing of the GOP, how are these voters different/similar to Republican voters in the past, in regards to the social context of just how different politics feel to compared how they did during the Obama and Bush administrations, for instance.

# Hypothesis: 

I believe that Donald Trump's rise to prominence ahead of the 2016 presidential election led to an increase in prejudiced views/values being held by United States citizens, as his rhetoric and overall demeanor fostered a resurgence in anti-progressive views.

# Why Did I choose this topic:

I believe that Donald Trump's rise as the leader of the Republican party stems from a multitude of complex social, economic, and political factors. That is not exactly what I seek to answer. We all know that Donald Trump came to the scene different than any other politicians that remain in our recent memories, harping on racial messaging and xenophobic ideas, Donald Trump has managed to engage with an entirely new section of the political community that in previous elections did not care much for politics. He has become the living light of American conservatism by combining these new tactics with the historical agenda of the GOP. Economic conservatism combined with cultural conservatism has allowed Trump's grip on middle America to be virtually unquestioned, as seen in his successes across the board in nearly every American county in the 2024 election. 

While Donald Trump has been generally careful to not say anything too outwardly offensive on a public platform, there is no doubt that his base and a large portion of his supporters carry such views and see Donald Trump as a protector of these forms of prejudice.

So, in the light of Trump's rise and his recent re-taking of the White House, has his, now, decade of political prominence resulted in an increase in hateful rhetoric and ideas. In this article, we will dive into measurements of support for LGBT+ families being able to adopt children. Has there been an increase in opposition to this basic right of an adult individual, or, in spite of Trump's volatility and perceived bigotry, is our society becoming more pluralistic?

Most of us still clearly remember when gay marriage was officially legalized under the Obama administration. This enactment was met with plenty of conservative backlash, so, now, over a decade later in 2024, do conservatives, and particularly Trump voters hold more, less, or similar prejudice towards the LGBT+ community than during Obama's terms.


Explanatory Variable:
- Before Trump era vs. After Trump Era
- This is because we are trying to find a correlation in the rise of anti-LGBT sentiment starting during and continuing after Trump's ascension.

Outcome Variable
-Is there more opposition to LGBT+ adoption?
-Increases in the proportion and magnitude of Americans in opposition to LGBT+ adoption since 2016 would indicate that my hypothesis, is indeed correct.

In other words, support for my hypothesis would be found if I do in fact find sufficient evidence that hate or negative feelings towards the LGBT+ community and its members' attempts to start families has risen since Trump first started captivating Americans by the millions.

## Project Milestone 3


```{r}
library(tidyverse)
library(readr)
library(dplyr)
anes_data <- read_csv("anes_condensed.csv")
```

Preliminary Data Wrangling

I will be utilizing the below three data frames to see if there is a rise in anti-gay sentiment among republican voters, before vs. after Trump's rise. 

Notes for renaming: VCF0004 is Year, VCF0302 IS Party Identification, VCF0878 is thoughts on gay adoption

# Filtration and Wrangling

```{r}
anes_republican <- anes_data |>
  filter(VCF0302 == 1) |>
  select(
    Year = VCF0004,
    Party = VCF0302,
    Gay_Adopt = VCF0878
  ) |>
  mutate(Era = ifelse(Year < 2016, "Pre Trump", "Trump-Era")) |>
  filter(Gay_Adopt != 8) |>
  mutate(Gay_Adopt = ifelse(Gay_Adopt == 1, "Support", "Oppose")) |>
  drop_na(Gay_Adopt)
```
# Visualization

```{r}
ggplot(anes_republican, aes( x = Gay_Adopt, fill = Gay_Adopt)) +
  geom_bar(color = "black") +
  facet_wrap(~Era) +
  #Separate colors for oppose and support
  scale_fill_manual(values = c("Oppose" = "indianred1", "Support" = "dodgerblue"))+
  labs(
    title = "Republican Voters' Sentiment on Gay/Lesbian Adoption of Children",
    subtitle = "Comparing the Pre and Post Trump Landscape",
    x = "Position on Gay/Lesbian Adoption Rights",
    y = "Number of Republican Respondents"
  ) +
  theme_minimal() +
  #remove legend
  theme(legend.position = "none")
```
Source: American National Election Studies (ANES)

## Project Milestone 4

#Second Visualization/Table

```{r}
library(knitr)
sum_up_table <- anes_republican |>
  group_by(Era, Gay_Adopt) |>
  summarize(Count = n())
kable(sum_up_table, col.names = c("Era", "Position on LGBT+ Adoption of Children", "Number of Respondents"),
      caption = "Pre and Post-Trump Sentiment on LGBT Child Adoption")
```
Source: American National Election Studies (ANES)

This data proves my hypothesis wrong. Although Donald Trump's presidency and future term have been highlighted by various instances of racial, cultural antagonism, support for LGBT+ adoption has actually increased. This is very interesting to me. A major concern for liberals this cycle was Trump and his base's desire to limit freedoms for the LGBT+ community, but clearly, the majority of voters support the security of basic liberties and the ability to form a family for these individuals. Now, I am led to believe that this increase is in part due to the successes of previous administrations and continued efforts to de-stigmatize previously marginalized members of society. For years now, gay marriage has been legal, and at this point, Americans would likely categorize it as just a fact-of-life. Even if one is opposed to same-sex relations, the legality of their love and marriage is secured. Although Trump's campaign this year often focused on sex changes for illegals and other volatile rhetoric attacking those who identify with the pronouns they/them in our society, most of his supporters believe in basic rights for all of these people. Instead, his voters are much more likely to be uncomfortable with aspects of LGBT+ life bleeding into their own. This, of course, is a problem. We need our society to be more accepting of people, regardless of race, sexual orientation, creed, etc. Yet, in spite of this, there is no significant evidence proving that Donald Trump has led an effort to outright oppress and subjugate the LGBT+ community. 